---
title: "Installing polySimIBD"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Installing polySimIBD}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Installing and Loading `polySimIBD`
To install this package, you will need to first install the [`remotes`](https://cran.r-project.org/web/packages/remotes/index.html) package and then run the following command:   

```{r, eval=FALSE}
remotes::install_github("nickbrazeau/polySimIBD")
```

### Dependencies
`polySimIBD` relies on the [Rcpp](https://cran.r-project.org/web/packages/Rcpp/index.html) package, which requires certain OS-specific dependencies: 

* Windows
    - Download and install the appropriate version of [Rtools](https://cran.rstudio.com/bin/windows/Rtools/) for your version of R. On installation, ensure you check the box to arrange your system PATH as recommended by Rtools
* Mac OS X
    - Download and install [XCode](http://itunes.apple.com/us/app/xcode/id497799835?mt=12)
    - Within XCode go to Preferences : Downloads and install the Command Line Tools
* Linux (Debian/Ubuntu)
    - Install the core software development utilities required for R package development as well as LaTeX by executing
    ```
    sudo apt-get install r-base-dev texlive-full
    ```


If you have any problems installing then please [raise an issue](https://github.com/nickbrazeau/polySimIBD/issues) on Github.

Assuming everything installed correctly, you can now load the package:

```{r, eval=FALSE}
library(polySimIBD)
```


Separately, after installing, if you identify any bugs, please [raise an issue](https://github.com/nickbrazeau/polySimIBD/issues) on Github so they can be addressed.
