---
output: html_document
editor_options: 
  chunk_output_type: console
---
---
title: "Extension of Model for Spatial Demes"
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{Extension of Model for Spatial Demes}
%\VignetteEngine{knitr::rmarkdown}
%\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(polySimIBD)
library(ggplot2)
library(tidygraph) # useful for manipulations & visaulizations
```

# Extending the Model for Space
We extend the discrete-loci, discrete-time structured Wright Fisher model for space by considering spatial _demes_ as a collection of hosts. Hosts move between demes based on a migration matrix *M*. In the spirit of the Wright-Fisher, the number of hosts within the deme (_i.e._ deme size) does not change over time. As before, , where each host is a "micro-deme" in order to model complexity of infection. 

## Forwards-in-time Description 
The full mathematical formulation of the nonspatial model can be found in the Supplementary Section of [Verity, Aydemic, Brazeau _et al._ Nat Comms 2020 (PMC7192906), Biorxiv](https://www.nature.com/articles/s41467-020-15779-8). The spatial extension to the model is described in [Brazeau _et al._ Biorxiv]().    

As a brief lay summary, we assume that each deme (_hereafter_ macrodeme) is a subpopulation within a very large population ($\iota \in \nu$, where $\iota$ is an individual macrodeme and $\nu$ is the total number of macrodemes in the entire population) and that each macrodeme represents a collection of individual hosts. Then, as before, let each individual host be represented by a deme  (_hereafter_ macrodeme), or a subpopulation within this large macrodeme ($i \in N_{\iota}$, where $i$ is an individual host and $N_{\iota}$ is the total host population for that macrodeme).   

We then allow the $J$ parasites (that reside within the host microdeme) to mate at random with the previous generation of parasites ($t_{1-}$) and produce a large number of parasite progeny. During mating, genetic recombination has the potential to occur based on the length of the genome and the recombination rate, $\rho$. Progeny are then allowed to move to a new macrodeme or stay in the current macrodeme based on the migration matrix, $\mathbf{M}$. Within the macrodemes, the progeny are then allowed migrate to a new or the same host microdeme with a probability of $\frac{m}{N_{\iota}}$. Progeny are then culled down to a smaller number of parasites per host by drawing from a Poisson distribution with a mean COI. In essence, the nonspatial and spatil simulator only differ by an additional hierarchical level of "demes" (macrodemes versus microdemes).


# Island Case
Here, we will make a dummy case of two island populations (population A vs. B) that are only marginally connected.   

Run the simulation:   
```{r}
# vectors must be ordered for population A and B
demesizes <- c(10, 15)
coimeans <- c(1.25, 1.25)
m <- c(0.25, 0.25)
migr_dist_mat <- matrix(0, ncol = 2, nrow = 2)
diag(migr_dist_mat) <- 100
swf <- polySimIBD::sim_swf(pos = sample(1:1e3, size = 50), 
                           N = demesizes, 
                           m = m, 
                           mean_coi = coimeans, 
                           migr_dist_mat = migr_dist_mat, 
                           rho = 1e-2, 
                           tlim = 10)
```
Pull out the ARG and then calculate the pairwise IBD.
```{r}
ARG <- polySimIBD::get_arg(swf)

# here we are going to get all the pairwise IBD combinations for our hosts
adj_graph <- t(combn(x = c(1:sum(demesizes)), m = 2)) %>% 
  tibble::as_tibble(., .name_repair = "minimal") %>% 
  magrittr::set_colnames(c("smpl1", "smpl2")) %>% 
  dplyr::mutate(pairwiseIBD = purrr::map2_dbl(.x = smpl1, .y = smpl2, .f = function(x,  y){
    polySimIBD:::quiet(
      polySimIBD::get_realized_pairwise_ibd(swf = swf, host_index = c(x,y))
    )
  }))

```

Let's visualize our simple scenarios
```{r}
adj_graph %>% 
  ggplot() + 
  geom_tile(aes(x = smpl1, y = smpl2, fill = pairwiseIBD)) + 
  scale_fill_viridis_c("Pairwise IBD") + 
  xlab("Sample 1") + ylab("Sample 2") + 
  labs(caption = "Samples 1:10 are on island A; Samples 11:25 are on island B") + 
  theme_minimal()

```
As can be seen from the figure, samples 1-10, which reside on island A have no relatedness to samples 11-25 that reside on island by B as they have been completely isolated through our migration matrix. 
