---
title: "_msprime_ and _polySimIBD_ Comparison"
output: 
  html_document:
    highlight: textmate
    theme: lumen
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, eval = T, results = 'hide',
                      fig.align = 'center', fig.width = 8, fig.height = 8)
```

## Model Formulation
As a sanity check, our `polySimIBD` model was tested against the `msprime` simulator. To make the simulator comparable, we considered `polySimIBD` with 1 deme (one indiviudal) with migration rate set to 1 (panmictic) and a mean COI of 1e3 (haploid). The `msprime` simulator was set to 5e2 indivduals (simulator params are in: `msprime_sims_forSWF.sh`). 

```{r}
library(tidyverse)

#....................................
# Read in msprimes and manipulate
#....................................
msprimesims <- readRDS("~/Documents/GitHub/polySimIBD/simulations/msprimesims/simdata/msprimesims.rds")
msprimesims <- unlist(msprimesims, recursive = F)

msprimesims$`discrete wright fisher.2` <- # overwrite t1 name for easier bind
  lapply(msprimesims$`discrete wright fisher.2`, function(x){
  colnames(x) <- c("t1")
  return(x)
})

msprimesims.df <- tibble(sample_size = c(2,3,5),
                         simsposexp = msprimesims)
msprimesims.df <- msprimesims.df %>%
  tidyr::unnest(cols = simsposexp)

# this craziness to get R purrr to cooperate
msprimesims.df$simsposexp <- purrr::map(msprimesims.df$simsposexp,
                                        tibble::as_tibble)

msprimesims.df <- msprimesims.df %>%
  tidyr::unnest(cols = simsposexp) %>%
  dplyr::mutate(model = "msprime")
msprimesims.df$pos <- 1:1001


#....................................
# Read in polysimibd
#....................................
polySimIBDsims.df <- readRDS("~/Documents/GitHub/polySimIBD/simulations/msprimesims/simdata/polySimIBD_sims.RDS") %>%
  dplyr::mutate(model = "polySimIBD") %>%
  dplyr::select(-c("intvl"))
polySimIBDsims.df$pos <- 1:1001


```


## Comparisons
### Continuous
Considering `msprime` with continuous sequencing/intervals. 
```{r}
#################################################################################
###################                Comparisons                ###################
##################################################################################

#....................................
# CONTINUOUS time for msprime
#....................................
simdf <- rbind.data.frame(msprimesims.df, polySimIBDsims.df)

simdf <- simdf %>%
  dplyr::select(c("sample_size", "pos", "model", dplyr::everything())) %>%
  tidyr::gather(., key = "Tn", value = "coaltime", 4:ncol(.)) %>%
  dplyr::filter(!is.na(coaltime)) 



# drop rows for reasonable plot
rows <- sample(1:nrow(simdf), size = 1e4)
rows <- c( 1:nrow(simdf) %in% sort(rows) )
simdfplotObj <- simdf %>%
  dplyr::filter(rows) %>%
  ggplot() +
  geom_density(aes(x = coaltime, y = stat(count),
                  fill = factor(model)), alpha = 0.5, color = NA) +
  facet_wrap(Tn~.) +
  xlab("Coalescence Time") +
  scale_fill_viridis_d("Model") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90),
        axis.title.y = element_blank())

jpeg("~/Documents/GitHub/polySimIBD/simulations/msprimesims/compare_polysim_mpsrime_cont.jpg", width = 11, height = 8, res = 300, units = "in")
plot(simdfplotObj)
graphics.off()



```

```{r, results='asis'}
plot(simdfplotObj)
```

```{r}

# drop rows for reasonable plot
rows <- sample(1:nrow(simdf), size = 1e4)
rows <- c( 1:nrow(simdf) %in% sort(rows) )
simdfplotObj <- simdf %>%
  dplyr::filter(rows) %>%
  ggplot() +
  ggridges::geom_density_ridges(aes(x = coaltime, y = factor(model),
                  fill = factor(model)), color = NA) +
  facet_wrap(Tn~.) +
  xlab("Coalescence Time") +
  scale_fill_viridis_d("Model") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, face = "bold"),
        axis.title.y = element_blank())

jpeg("~/Documents/GitHub/polySimIBD/simulations/msprimesims/compare_polysim_mpsrime_cont_ggridges.jpg", width = 11, height = 8, res = 300, units = "in")
plot(simdfplotObj)
graphics.off()

```

```{r, results='asis'}
plot(simdfplotObj)
```




```{r, results='asis'}

simdf.summary <- simdf %>%
  dplyr::group_by(model, Tn) %>%
  dplyr::summarise(
    minTime = min(coaltime),
    meanTime = mean(coaltime),
    medTime = median(coaltime),
    maxTime = max(coaltime),
    sdTime = sd(coaltime)
  ) %>%
  dplyr::arrange(., Tn)

simdf.summary %>% 
  kableExtra::kable(format = "html", escape = F) %>%
  kableExtra::kable_styling("striped", full_width = F) 


```



### Discrete
Considering `msprime` with discretized sequencing/loci. 

```{r}


#....................................
# DISCRETE time for msprime
#....................................

# find discrete rows
discretepos <- seq(1, 1001, by = 100)
discreterows <- which(simdf$pos %in% discretepos)
discreterows <- c(1:nrow(simdf) %in% discreterows)



simdfplotObj <- simdf %>%
  dplyr::filter(discreterows) %>%
  ggplot() +
  geom_density(aes(x = coaltime, y = stat(count),
                  fill = factor(model)), color = NA, alpha = 0.5) +
  facet_wrap(Tn~.) +
  xlab("Coalescence Time") +
  scale_fill_viridis_d("Model") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90),
        axis.title.y = element_blank())



jpeg("~/Documents/GitHub/polySimIBD/simulations/msprimesims/compare_polysim_mpsrime_discrete.jpg", width = 11, height = 8, res = 300, units = "in")
plot(simdfplotObj)
graphics.off()



```


```{r, results='asis'}
plot(simdfplotObj)
```


```{r}


# drop rows for reasonable plot
rows <- sample(1:nrow(simdf), size = 1e4)
rows <- c( 1:nrow(simdf) %in% sort(rows) )
simdfplotObj <- simdf %>%
  dplyr::filter(rows) %>%
  ggplot() +
  ggridges::geom_density_ridges(aes(x = coaltime, y = factor(model),
                  fill = factor(model)), color = NA) +
  facet_wrap(Tn~.) +
  xlab("Coalescence Time") +
  scale_fill_viridis_d("Model") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, face = "bold"),
        axis.title.y = element_blank())

jpeg("~/Documents/GitHub/polySimIBD/simulations/msprimesims/compare_polysim_mpsrime_disc_ggridges.jpg", width = 11, height = 8, res = 300, units = "in")
plot(simdfplotObj)
graphics.off()

```

```{r, results='asis'}
plot(simdfplotObj)
```



```{r, results='asis'}

simdf.summary <- simdf %>%
  dplyr::filter(discreterows) %>%
  dplyr::group_by(model, Tn) %>%
  dplyr::summarise(
    minTime = min(coaltime),
    meanTime = mean(coaltime),
    medTime = median(coaltime),
    maxTime = max(coaltime),
    sdTime = sd(coaltime)
  ) %>%
  dplyr::arrange(., Tn)

simdf.summary %>% 
  kableExtra::kable(format = "html", escape = F) %>%
  kableExtra::kable_styling("striped", full_width = F) 
```





